Hydras \& Co.

### 2022-05-06 doc/gaia-chapter.tex
minor changes in theories/gaia
minor corrections

The bridge we aim to build and cross is a two-way bridge:

\begin{itemize}
\item \gaia contains lemmas on ordinal arithmetic (\emph{e.g.} associativity of multiplication of ordinals strictly less than $\epsilon_0$), which were missing from \Hydras. 
  \item The other way, \Hydras contains formal definitions and proofs of properties of canonical sequences and rapidly growing functions  (see Chapters~\ref{chap:ketonen} and~\ref{chap:alpha-large}~of this book).
  \end{itemize}

  We plan to make accessible the first kind of lemmas to \HydrasLib' users and the second kind to \gaia's. For this purpose, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
  Please note that by ``import'' we do not mean ``duplicate''
  nor ``repairing'' \HydrasLib' proofs.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs. In a future version, it would be nice to have our proofs in the same context, after porting \HydrasLib' proofs under \gaia's
  definition, following  proof re-use or repair  techniques~\cite{Barthe2001, Magaud03a, ringer2021}.

@@

Many constants are common to \HydrasLib and \gaia. For instance, the name \texttt{zero} may refer to \gaia's
\texttt{gaia.ssete9.CantorOrdinal.zero}  and
\HydrasLib' \texttt{hydras.Epsilon0.T1.zero}. By default, the former will be simply called \texttt{zero} and the latter \texttt{T1.zero}.

@@

\subsubsection{A few constants}
For each constant: $0$, $1$, \dots, $n$ and $\omega$, we prove
that \texttt{hydras}' definition refines \texttt{gaia}'s.
\inputsnippets{T1Bridge/constantRefs}

@@

\subsubsection{Order and comparison}
\label{sect:lt-compat-gaia}
Leibniz equality and strict order on both libraries are
compatible:

@@

The binary operations $+$ and $\times$ are defined the same way in both libraries. Please note that they require comparison of ordinal terms. So, the proof of the following lemmas applies
compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).

@@

\subsubsection{Cantor normal form}
\label{nf-gaia-compat}
Both definitions of ``being in Cantor normal form'' are
compatible.

@@

\subsection{Hessenberg sum}
\index{gaiabridge}{Hessenberg sum}
Natural sum  (a.k.a. Hessenberg  sum) is defined in
\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).
We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.

@@

\inputsnippets{GHessenberg/oplusNf}
\inputsnippets{GHessenberg/oplusEquations}
\inputsnippets{GHessenberg/oplusEquationsb}
\inputsnippets{GHessenberg/oplusEquationsc}

@@

\inputsnippets{GPaths/KSThm24}
\inputsnippets{GPaths/Cor12}

@@

We plan to automatize as soon as possible this kind of transfer of lemmas from \gaia to \HydrasLib.

>>>>

The main objective of the \texttt{theories/gaia/*.v} files is
to provide \gaia's users definitions and lemmas on
some combinatorial aspects of ordinal numbers less than $\epsilon_0$ \cite{KS81}, already proven in \Hydras: canonical sequences, accessibility properties, large sets and rapidly growing arithmetical functions.
We may also complete \Hydras with some lemmas proven in \gaia.

For these purposes, we write modules dedicated to the importation of definitions and lemmas from each of both libraries into the other one.
  Please note that by ``import'' we do not mean ``duplicate''
  nor ``repairing'' \HydrasLib' proofs.  Whenever possible, we proceed by rewriting steps over the statements and not the proofs. In a future version, it would be nice to have our proofs in the same context, after porting \HydrasLib' proofs under \gaia's
  definition, following  proof re-use or repair  techniques~\cite{Barthe2001, Magaud03a, ringer2021}.

@@

Likewise, many constants are common to \HydrasLib and \gaia. For instance, the name \texttt{zero} may refer to \gaia's
\texttt{gaia.ssete9.CantorOrdinal.zero}  and
\HydrasLib' \texttt{hydras.Epsilon0.T1.zero}. By default, the former will be simply called \texttt{zero} and the latter \texttt{T1.zero}.

@@

\subsubsection{A few constants}
For each constant: $0$, $1$,\dots, $n$ and $\omega$, we prove
that \texttt{hydras}' definition refines \texttt{gaia}'s.
\inputsnippets{T1Bridge/constantRefs}

@@

\subsubsection{Order and comparison}
\label{sect:lt-compat-gaia}
The  strict order on both types \texttt{T1} are
compatible:

@@

The definition of the binary operations $+$ and $\times$
requires the comparison of ordinal terms, which is not implemented the same way in both libraries. Thus, the proof of the following lemmas applies applies
compatibility lemmas like \texttt{compare\_ref} (see Section~\vref{sect:gaia-compare-ref}).

@@

\subsubsection{Well formed ordinal terms (Cantor normal form)}
\label{nf-gaia-compat}
Both definitions of ``being in Cantor normal form'' are
compatible.

@@

\subsection{Hessenberg sum}
\index{gaiabridge}{Hessenberg sum}
Natural sum  (a.k.a. Hessenberg  sum) is defined in
\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).

Instead of defining this operation \emph{ex nihilo} in \texttt{theories/gaia}, we chose to \emph{import} the definition
from \href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} as follows:

@@

\inputsnippets{GHessenberg/oplusEquationsc}
\inputsnippets{GHessenberg/oplusEquations}
\inputsnippets{GHessenberg/oplusEquationsb}
\inputsnippets{GHessenberg/oplusNf}

@@

%\inputsnippets{GPaths/KSThm24}
\inputsnippets{GPaths/Cor12}

@@

We plan to automatize as soon as possible the proof of this kind of transfer lemmas from \gaia to \HydrasLib.
